Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1) \note{\scriptsize Authors, please add/check middle initials and affiliations}

### 2014-04-12 WSSSPE1_report/WSSSPE1_summary_report.tex
Section 1. Add non-breaking spaces. A minor change.

Because progress in scientific research is dependent on the quality of and
accessibility to software at all levels, it is now critical to address many
challenges related to the development, deployment, and maintenance of reusable
software.
%\note{Ketan: What are the levels in scientific process where software
%is required. It might be interesting to outline these stages and types of
%software required in each stage.}
In addition, it is essential that scientists,
researchers, and students are able to learn and adopt software-related skills
and methodologies. Established researchers are already acquiring some of these
skills, and in particular a specialized class of software developers is
emerging in academic environments as an integral part of
successful research teams. The WSSSPE workshop was intended to provide a forum
for discussion of the challenges, including both positions and experiences. The
short papers and debates have been archived to provide a basis for continued
discussion, and the workshop has fed into the collaborative writing of this
document. An estimate of 90 to 150 participants were present
throughout the day. Additional papers based on extended versions of workshop
submissions are expected. The level of interest in the workshop has led the
organizers, working with some of the submitters and attendees, to plan two
additional workshops (at the 2014 SciPy and SC14 conferences), to obtain
funds from the National Science Foundation and the Gordon and Betty Moore
Foundation to support these workshops, and to turn the
workshop website into a community website that can be used as a focus for
further discussion and progress. Additionally, a minisymposium that aims to
further explore some of the key issues raised in WSSSPE is co-organized by a
WSSSPE1 participant at the 2014 Society for Industrial and Applied Mathematics
(SIAM) Annual Meeting on ``Reliable
Computational Science.''

The goal of this paper is to summarize the contents of the workshop papers
and discussion, and identify issues that future WSSSPE events will address.
Before the WSSSPE1 workshop took place, the organizers self-published
a paper~\cite{WSSSPE1-pre-report} to document the process of
organizing and advertising the workshop, collecting and reviewing the
papers, and putting together the agenda. Section~\ref{sec:process} is
based on that publication. The remainder of this paper is based on the
workshop itself, as documented in a set of collaborative notes taken
during the workshop~\cite{WSSSPE1-google-notes}, including discussion
about the keynote presentations (\S\ref{sec:keynotes}), the three
panels (\S\ref{sec:devel}-\ref{sec:community}), 
cross-cutting issues (\S\ref{sec:cross-cutting}),
use cases (\S\ref{sec:use-cases}), and conclusions
(\S\ref{sec:conclusions}).  Additionally, the accepted workshop papers are listed in
Appendix~\ref{sec:papers}, and a partial list of workshop attendees is
in Appendix~\ref{sec:attendees}.

>>>>

Because progress in scientific research is dependent on the quality of and
accessibility to software at all levels, it is now critical to address many
challenges related to the development, deployment, and maintenance of reusable
software.
%\note{Ketan: What are the levels in scientific process where software
%is required. It might be interesting to outline these stages and types of
%software required in each stage.}
In addition, it is essential that scientists,
researchers, and students are able to learn and adopt software-related skills
and methodologies. Established researchers are already acquiring some of these
skills, and in particular a specialized class of software developers is
emerging in academic environments as an integral part of
successful research teams. The WSSSPE workshop was intended to provide a forum
for discussion of the challenges, including both positions and experiences. The
short papers and debates have been archived to provide a basis for continued
discussion, and the workshop has fed into the collaborative writing of this
document. Estimates of 90 to 150 participants were present
throughout the day. Additional papers based on extended versions of workshop
submissions are expected. The level of interest in the workshop has led the
organizers, working with some of the submitters and attendees, to plan two
additional workshops (at the 2014 SciPy and SC14 conferences), to obtain
funds from the National Science Foundation and the Gordon and Betty Moore
Foundation to support these workshops, and to turn the
workshop website into a community website that can be used as a focus for
further discussion and progress. Additionally, a minisymposium that aims to
further explore some of the key issues raised in WSSSPE is co-organized by a
WSSSPE1 participant at the 2014 Society for Industrial and Applied Mathematics
(SIAM) Annual Meeting on ``Reliable
Computational Science.''

The goal of this paper is to summarize the contents of the workshop papers
and discussion, and identify issues that future WSSSPE events will address.
Before the WSSSPE1 workshop took place, the organizers self-published
a paper~\cite{WSSSPE1-pre-report} to document the process of
organizing and advertising the workshop, collecting and reviewing the
papers, and putting together the agenda. Section~\ref{sec:process} is
based on that publication. The remainder of this paper is based on the
workshop itself, as documented in a set of collaborative notes taken
during the workshop~\cite{WSSSPE1-google-notes}, including discussion
about the keynote presentations~(\S\ref{sec:keynotes}), the three
panels~(\S\ref{sec:devel}-\ref{sec:community}), 
cross-cutting issues~(\S\ref{sec:cross-cutting}),
use cases~(\S\ref{sec:use-cases}), and conclusions~(\S\ref{sec:conclusions}).  
Additionally, the accepted workshop papers are listed in
Appendix~\ref{sec:papers}, and a partial list of workshop attendees is
in Appendix~\ref{sec:attendees}.
